---
title: "Leaflet_tester"
date: 06-30-2023
categories: [Data-Viz, openxlsx, stringr, fuzzyjoin, mapview, sf] # self-defined categories
draft: false # setting this to `true` will prevent your post from appearing on your listing page until you're ready!
---

This is a draft of a project for my father. I want to make sure the photo files structure renders properly online. This will be revised and reposted later as a tutorial. It will also be deleted once I get his input.

```{r}
library(tidyverse) # who doesn't want to be tidy?
library(ggthemes) # more themes for ggplot
library(gt) # For nice tables
library(ggrepel) # to help position labels in ggplot graphs
library(openxlsx) # importing excel files from a URL
library(fuzzyjoin) # for joining on inexact matches
library(sf) # for handling geo data
library(mapview) # quick interactive mapping
library(leaflet) # more mapping
library(here)
library(magick)

```

Reading in his spreadsheet

```{r}
blog_folder <- "posts/2023-06-30-leaflet2"
photo_folder <- "Photos"
archive_folder <- "Archive"
problem_folder <- "Problem"
data_file <- "Tombstone_Data_small.xlsx"
#data_file <- "Tombstone Data.xlsx"
# read in excel sheet
tombstones <-
  read.xlsx(here(blog_folder, data_file),
    sheet = 1
  )

```

Cleaning it up.

1.  Change NAs in middle name variable to "".

```{r}
tombstones <- tombstones %>%
mutate(Middle.Name = ifelse(is.na(Middle.Name), "", Middle.Name)) %>%
  mutate(Cemetery = ifelse(is.na(Cemetery), "", Cemetery))

```

2.  Latitude and longitude data contains some stray degree and minute symbols. The degree symbol appears both as a straight and curved apostrophe and the degree symbols appear both as o and O. This cleaning needs to be done on both N and W columns.

```{r}

tombstones <- tombstones %>%
mutate(N = str_replace_all(N, pattern = "’", " ")) %>%
mutate(N = str_replace_all(N, pattern = "O", " "))  %>%
mutate(N = str_replace_all(N, pattern = "o", " "))  %>%
mutate(N = str_replace_all(N, pattern = "'", " "))  %>% 
mutate(W = str_replace_all(W, pattern = "’", " ")) %>%
mutate(W = str_replace_all(W, pattern = "O", " "))  %>% 
mutate(W = str_replace_all(W, pattern = "o", " ")) %>%
mutate(W = str_replace_all(W, pattern = "'", " "))  
```

Next, I'm converting the N and W data to decimal latitude and longitude. S/W should be "-" and N/E should be "+". I split the degree/minute/second data into parts and then do the conversion. I delete the intermediate components when done.

```{r}

#converting to decimal latitude
tombstones <- tombstones %>%
  mutate(part1N = str_split_fixed(N, pattern = " ", n = 2) ) %>%
  mutate(N_degree = as.numeric(part1N[,1])) %>%
  mutate(part2N = str_split_fixed(part1N[,2], pattern = '\\.', n = 2)) %>%
  mutate(N_minute = as.numeric(part2N[,1])) %>%
  mutate(N_second = as.numeric(part2N[,2])) %>%
  mutate(lat = N_degree + N_minute/60 + N_second/3600)

#converting to decimal longitude  
tombstones <- tombstones %>%
  mutate(part1W = str_split_fixed(W, pattern = " ", n = 2) ) %>%
  mutate(W_degree = as.numeric(part1W[,1])) %>%
  mutate(part2W = str_split_fixed(part1W[,2], pattern = '\\.', n = 2)) %>%
  mutate(W_minute = as.numeric(part2W[,1])) %>%
  mutate(W_second = as.numeric(part2W[,2])) %>%
  mutate(long = -(W_degree + W_minute/60 + W_second/3600)) 

tombstones <- tombstones %>%
  select(-contains("part"))


```

My father's photo naming convention is "last name first name middle initial". I'm going to construct a variable full_name from the spreadsheet data and use that to match. Initially, I'm leaving out middle name, because that column is problematic.

```{r}
tombstones <- tombstones %>%
  mutate(full_name = paste(Surname, First.Name, sep = " "))
```

Now, I'm generating a list of all the photo names and converting it to a dataframe. I'm going to match against this list.

```{r}
photo_names = list.files(here(blog_folder, photo_folder))
photo_df = as.data.frame(photo_names)


```

Doing a fuzzy join on the two dataframes.

```{r}
tombstones_merged <- fuzzy_right_join(photo_df, tombstones, 
                             by = c("photo_names" = "full_name"),
                             match_fun = str_detect)

```

Seeing what photos weren't matched. I'm actually going to move them somewhere else.

```{r}
not_matched <- anti_join(photo_df, tombstones_merged)

index <- 1
for (index in seq(1:nrow(not_matched))) {
  file.rename(from = here(blog_folder, photo_folder, not_matched$photo_names[index]),
                         to = here(blog_folder, problem_folder, not_matched$photo_names[index] ))
}


```

I'm dropping all the NAs in the photo_names column. I'm creating an image tag, because I'd ideally like to use that for the photos.

```{r}

# turn photo tag
tombstones_merged <- tombstones_merged %>%
  drop_na(photo_names) %>%
  mutate(tag = paste0("<img src=", here(blog_folder, photo_folder, photo_names),">"))

```

How many people have multiple photos?

```{r}
multiples <- tombstones_merged %>% group_by(full_name) %>% count(sort = TRUE) %>% filter(n > 1)
```

Look at these multiples more closely

```{r}

df_temp <- tombstones_merged %>%
  filter(full_name == multiples$full_name[1]) 
df_temp

```

Write a function to handle making panel figures. Not a function yet

https://stackoverflow.com/questions/61196196/image-append-on-dynamic-number-of-variables

```{r}

#testing
if (nrow(multiples) > 0 ) {
index <- 1

for (index in seq(1:nrow(multiples))) {
  df_temp <- tombstones_merged %>%
    filter(full_name == multiples$full_name[index])
  df_temp
  these <-
    as.list(here(blog_folder, photo_folder, df_temp$photo_names))
  photo_panel <-
    image_append(do.call("c", lapply(these, image_read)))
  image_write(
    photo_panel,
    path =  here(
      blog_folder,
      photo_folder,
      paste0(df_temp$full_name[1], "_panel.png")
    ),
    format = "png"
  )
  index2 <- 1
  for (index2 in seq(1:nrow(df_temp))) {
    file.rename(
      from = here(blog_folder, photo_folder, df_temp$photo_names[index2]),
      to = here(blog_folder, archive_folder, df_temp$photo_names[index2])
    )
  }
}
}
```

Now I need to redo everything so it will only match on the panel figures and not the ones I moved.

```{r}
photo_names2 = list.files(here(blog_folder, photo_folder))
photo_df_final = as.data.frame(photo_names2)
tombstones_final <- fuzzy_right_join(photo_df_final, tombstones, 
                             by = c("photo_names2" = "full_name"),
                             match_fun = str_detect)
```

Dropping any NAs in lat/long. Converting it to a SF geo object. I'm adding some jitter to the coordinates; tombstones that are very close to each other have the same coordinates and show up as overlapping on the map.

```{r}
tombstones_final <- tombstones_final %>% drop_na(lat) %>% drop_na(long)


tombstones_geo <- st_as_sf(tombstones_merged, coords = c("long", "lat"), crs = 4326)
tombstones_geo <- st_jitter(tombstones_geo, factor = 0.001)


```

Final map.

```{r}
#| column: page
#| 
image_list <- tombstones_geo$photo_names

labels <- sprintf(
  "<strong>tombstones_geo$full_name</strong> <br/>,
  tombstones_geo$Cemetery" ) %>% lapply(htmltools::HTML)

leaflet() %>%
  addTiles() %>%
  #addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    data = tombstones_geo,
    label = ~ (paste( First.Name, Surname, "-", Cemetery, "Cemetery", sep = " ")), 
   # label = ~labels,
   # popup = leafpop::popupImage(paste0(here(blog_folder, photo_folder),"/", image_list)),
    clusterOptions = markerClusterOptions(),
    opacity = 1,
    radius = 5,
    color = "black",
    stroke = NA,
    group = "group1"
  )  %>%

# popup = ~tag,


   leafpop::addPopupImages(
    image = paste0(here(blog_folder, photo_folder),"/", image_list),
     src = local,
     group = "group1", width = 600,
    maxHeight = 300, maxWidth = 600
   )

```
